---
title       : Machine memory requirements
subtitle    : 
author      : Anne Troop
job         : WebSphere Admin
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Our Initial Problem



- Applications hosted on virtual machines
- Initial estimate of how much memory we needed per hosted application was somewhat high
- Wanted data to narrow down what the requirements really are

--- .class #id 


## Action


- We decided to collect data on memory in active use
- Using memstat
- Output looks like this:


2015-01-01 00:00:01 18312 11779

2015-01-01 00:01:01 18360 11730

2015-01-01 00:02:01 18414 11677

2015-01-01 00:03:01 18431 11659

2015-01-01 00:04:01 18433 11657

2015-01-01 00:05:01 18436 11654

...



---


## Data

```{r, cache=FALSE, echo=FALSE}


      library(plyr)
      library(lubridate)
      library(lattice)
      
      dir <- "assets/data"
      #filenames <- list.files(dir, pattern="*.out", full.names=TRUE)

      listdf <- rep(NULL, 3)
      filename <- rep(NULL, 3)
      for (i in 1:3) {
        filename[i] <- paste(dir, "/machine",i,"_memusage.out", sep="");
        listdf[[i]] <- read.table(filename[i])
        listdf[[i]]$machine <- paste("machine",i,sep="");
      }


      mem <- rbind(listdf[[1]], listdf[[2]])
      mem <- rbind(mem, listdf[[3]])
      
        mem$when <- ymd_hms(paste(mem$V1, mem$V2))
        
        mem <- transform(mem, machine = factor(machine))     
        g <- ggplot(mem, aes(when, V3) ) 
        g <- g + geom_point(aes(color = machine))
        g <- g + facet_grid( . ~ machine)
        plot(g)

```




---


## Results



1. Maximum memory used was lower than expected
2. Able to reclaim resources and save money
3. Reports tell which machine has head-room to put another app on

---


## Amusing side effects


- Other data patterns noted were:
*  characteristic down-and-then-up of restarting an application
*  very large dips identify when the machines were rebooted

```{r, cache=FALSE, echo=FALSE}


      library(plyr)
      library(lubridate)
      library(lattice)
      
      dir <- "assets/data"
      #filenames <- list.files(dir, pattern="*.out", full.names=TRUE)

      listdf <- rep(NULL, 3)
      filename <- rep(NULL, 3)
      for (i in 1:3) {
        filename[i] <- paste(dir, "/machine",i,"_memusage.out", sep="");
        listdf[[i]] <- read.table(filename[i])
        listdf[[i]]$machine <- paste("machine",i,sep="");
      }


      mem <- rbind(listdf[[1]], listdf[[2]])
      mem <- rbind(mem, listdf[[3]])      
      mem$when <- ymd_hms(paste(mem$V1, mem$V2))
        
      mem <- transform(mem, machine = factor(machine))     

      g <- ggplot(mem, aes(when, V3) ) 
      g <- g + geom_point(aes(color = machine))
      g <- g + facet_grid( machine ~ .)

      plot(g)

```




